% ============================================================
%  MODULE X — Advanced Topics & Next Steps
% ============================================================

\chapter{Advanced Topics \& Next Steps}

\begin{notebox}
This module introduces areas to explore as you develop mastery of the foundations.
Each topic below is a deep discipline in its own right. Treat these as
signposts for your learning roadmap, not immediate requirements.
\end{notebox}

% ────────────────────────────────────────────────────────────
\section{Options — The Basics}

An \textbf{option} is a contract giving the holder the \textbf{right, but not the obligation},
to buy or sell an asset at a specified \textbf{strike price} by or on the \textbf{expiration date}.

\begin{center}
\begin{tabular}{>{\bfseries}p{2.5cm} >{\bfseries}p{2.5cm} p{7.5cm}}
\toprule
Type & Position & Rights and Risk \\
\midrule
Call & Long & Right to \emph{buy} at strike. Profit if price rises above strike + premium. \\
Call & Short & Obligation to \emph{sell} at strike. Limited profit (premium), unlimited risk. \\
Put & Long & Right to \emph{sell} at strike. Profit if price falls below strike + premium. \\
Put & Short & Obligation to \emph{buy} at strike. Limited profit (premium), large downside. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{The Greeks}
Options price sensitivity measures:

\begin{center}
\begin{tabular}{>{\bfseries}p{2cm} p{10.5cm}}
\toprule
Greek & What It Measures \\
\midrule
Delta ($\Delta$) & Change in option price per \$1 change in underlying. Ranges 0--1 (calls), 0 to $-$1 (puts). \\
Gamma ($\Gamma$) & Rate of change of Delta. High near expiration. \\
Theta ($\Theta$) & Time decay. Options lose value as expiration approaches. \\
Vega ($\mathcal{V}$) & Sensitivity to changes in implied volatility. \\
Rho ($\rho$) & Sensitivity to interest rate changes. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Basic Strategies}
\begin{itemize}
  \item \textbf{Covered Call} — own stock + sell call = generate income, cap upside
  \item \textbf{Protective Put} — own stock + buy put = insure downside
  \item \textbf{Straddle} — buy call + buy put at same strike = profit from big move either way
  \item \textbf{Vertical Spread} — buy one option + sell another at different strike = defined risk/reward
\end{itemize}

\begin{warningbox}
Options involve leverage and time decay. Short options have unlimited or very large risk.
Understand all mechanics before trading with real capital.
\end{warningbox}

% ────────────────────────────────────────────────────────────
\section{Futures Contracts}

A \textbf{futures contract} is a \textbf{standardised agreement} to buy or sell an asset
at a predetermined price on a specific future date.

Uses:
\begin{itemize}
  \item \textbf{Hedging} — lock in a price (e.g., airline hedging jet fuel)
  \item \textbf{Speculation} — directional bets with leverage
  \item \textbf{Arbitrage} — exploit price discrepancies between spot and futures
\end{itemize}

Key futures markets: \textbf{equity index} (S\&P 500 /ES, NASDAQ /NQ),
\textbf{bond} (US Treasury /ZB, /ZN), \textbf{energy} (Crude /CL, Natural Gas /NG),
\textbf{metals} (Gold /GC, Silver /SI), \textbf{currencies} (/6E, /6J).

\subsection{Contango and Backwardation}
\begin{itemize}
  \item \textbf{Contango}: Futures price $>$ spot price. Normal for storage-cost commodities.
        Rolling long positions forward incurs a cost.
  \item \textbf{Backwardation}: Futures price $<$ spot price. Indicates supply tightness.
        Rolling forward generates a positive roll yield.
\end{itemize}

% ────────────────────────────────────────────────────────────
% Term Cards / Glossary
% ────────────────────────────────────────────────────────────
\section{Trading Term Cards (Finance Jargon)}

\subsection{Assets and Cash Markets}

\begin{keyterm}{Equities}
Stocks / shares that represent ownership in a company. If the company grows, the equity
value can rise; if it deteriorates, it can fall.
\\\textbf{Example:} Buying 1 share of a tech company makes you a partial owner of its
future cash flows (via dividends and/or resale value).
\end{keyterm}

\begin{keyterm}{Fixed Income}
Securities where returns are driven primarily by interest payments and credit quality
(for example: government or corporate bonds). You earn coupons, and your total return
depends on yields moving after you buy.
\\\textbf{Example:} Buy a 5\% coupon bond; if market yields drop, the bond price can rise
even if the coupon stays the same.
\end{keyterm}

\begin{keyterm}{FX (Foreign Exchange)}
Trading currencies in pairs (e.g., \textbf{EUR/USD}). An FX quote is just the price of one
currency in terms of another.
\\\textbf{Example:} If EUR/USD moves from 1.1000 to 1.1050, EUR strengthened versus USD.
\end{keyterm}

\begin{keyterm}{Commodities}
Raw materials (oil, gold, wheat, copper, etc.) traded directly (spot) or via derivatives
(futures/options). Many commodities are priced in USD, so FX and macro matter a lot.
\\\textbf{Example:} Oil prices can jump on supply disruptions; gold often responds to real
rates and risk sentiment.
\end{keyterm}

\subsection{Derivatives}

\begin{keyterm}{Futures}
Exchange-traded, standardised contracts to buy or sell an underlying at a future date
for a predetermined price. Futures are marked-to-market (gains/losses flow daily).
\\\textbf{Example:} Hedge jet fuel costs for next quarter by locking in the fuel price
via futures.
\end{keyterm}

\begin{keyterm}{Options}
Contracts that give the holder the \textbf{right} (not obligation) to buy (call) or sell
(put) the underlying at a \textbf{strike} price on or before expiration.
\\\textbf{Example:} A call option benefits if the stock rises above strike; your downside
is limited to the premium you paid.
\end{keyterm}

\begin{keyterm}{Perp Swaps (Perpetual Futures Swaps)}
Derivative contracts with no fixed expiration date (``perpetual''). Traders typically
use them like futures, but the contract uses \textbf{funding} to keep the perp price close
to the relevant spot/mark price.
\\\textbf{Example:} If BTC perp trades too high versus BTC spot, funding tends to push the perp
back down by incentivizing more short demand.
\end{keyterm}

\begin{keyterm}{Funding Rate in Perp Swaps}
A periodic payment exchanged between longs and shorts. The sign/magnitude is designed to
reduce persistent price gaps (basis) between perp and spot.
\\\textbf{Example:} If perp price trades above spot (positive premium), longs typically pay
shorts (funding is positive for longs), discouraging over-longing.
\end{keyterm}

\begin{keyterm}{Forwards}
OTC (custom) contracts agreeing to buy/sell an asset at a future date for a fixed price.
Forwards are not standardized like exchange futures and therefore typically carry more
\textbf{counterparty risk}.
\\\textbf{Example:} A producer and airline might agree on fixed fuel pricing for next year
via an OTC forward.
\end{keyterm}

\subsection{Position Direction and Portfolio Risk}

\begin{keyterm}{Long}
Holding a position that benefits when the underlying price rises.
\\\textbf{Example:} Buying BTC at \$60{,}000 and selling at \$63{,}000 profits if your entry is lower
than your exit.
\end{keyterm}

\begin{keyterm}{Short}
Holding a position that benefits when the underlying price falls (via borrowing/selling
or via derivatives like futures/perps).
\\\textbf{Example:} If you short at \$100 and it falls to \$95, you profit (before fees/interest).
\end{keyterm}

\begin{keyterm}{Flat}
No active exposure: your net position is zero (or close enough that price moves do not
meaningfully affect your P\&L).
\\\textbf{Example:} After closing a long and a short, you are flat and stop taking directional risk.
\end{keyterm}

\begin{keyterm}{Exposure}
How sensitive your portfolio is to a specific risk factor (directional price moves, rates, FX,
volatility, etc.). Think of exposure as ``how much you win/lose if the variable moves.''
\\\textbf{Example:} Long 1 BTC means your mark-to-market moves roughly with BTC price by about 1:1
(ignoring funding/fees).
\end{keyterm}

\begin{keyterm}{Margin}
Collateral you post to support leveraged positions. Margin determines how much loss you
can absorb before a forced close.
\\\textbf{Example:} With \$1{,}000 initial margin, you might control \$10{,}000 notional (10x),
but you only have \$1{,}000 of buffer.
\end{keyterm}

\begin{keyterm}{Leverage}
Controlling a larger notional position relative to your posted equity/collateral.
Leverage amplifies both gains and losses.
\\\textbf{Example:} 10x leverage means a 1\% adverse move can cost about 10\% of your margin
(ignoring fees and mark-price mechanics).
\end{keyterm}

\begin{keyterm}{Liquidation}
The forced closing of a leveraged position when collateral/equity falls below the
maintenance requirement (or the exchange's liquidation threshold).
\\\textbf{Example:} BTC drops quickly; your margin cushion is depleted and the exchange closes you
at the prevailing mark/liquidation price.
\end{keyterm}

\begin{keyterm}{PnL (Profit and Loss)}
The amount you gain or lose from a trade. It can be realized (closed positions) or
unrealized (mark-to-market on open positions).
\\\textbf{Example:} Buy at \$100 and sell at \$110: realized P\&L is +\$10 per unit.
\end{keyterm}

\begin{keyterm}{Sharpe Ratio}
A performance metric that rewards higher returns while penalizing volatility:
\[
S = \frac{R_p - R_f}{\sigma_p}
\]
where $R_p$ = portfolio return, $R_f$ = risk-free rate, and $\sigma_p$ = return volatility.
\\\textbf{Example:} Two strategies both return 10\%; if one is much smoother (lower volatility),
it tends to have a higher Sharpe.
\end{keyterm}

\begin{keyterm}{Counterparty Risk}
Risk that the other side in a contract fails to perform (e.g., defaults on payments).
Exchange-cleared derivatives often reduce this risk compared to OTC contracts.
\\\textbf{Example:} An OTC forward depends on your counterparty paying/settling correctly.
\end{keyterm}

\subsection{Order Book, Execution, and Trading Costs}

\begin{keyterm}{Bid}
The highest price a buyer is willing to pay.
\\\textbf{Example:} If the bid is \$99.90, buyers are currently willing to buy at \$99.90.
\end{keyterm}

\begin{keyterm}{Ask}
The lowest price a seller is willing to accept.
\\\textbf{Example:} If the ask is \$100.10, sellers are currently offering to sell at \$100.10.
\end{keyterm}

\begin{keyterm}{Spread}
The difference between ask and bid. In liquid markets, the spread is your baseline trading
cost (you usually pay it when you enter/exit).
\\\textbf{Example:} Bid \$99.90, ask \$100.10: spread is \$0.20.
\end{keyterm}

\begin{keyterm}{Fill}
When an order is executed. Orders can fill partially (some quantity filled) or fully.
\\\textbf{Example:} Your \$10{,}000 limit buy might fill \$6{,}000 if remaining book liquidity
doesn't reach your price.
\end{keyterm}

\begin{keyterm}{Slippage}
The difference between the price you expected and the actual execution price, driven by
spread, fast market moves, and your order size/impact.
\\\textbf{Example:} You expect a fill near \$100 but the market moves and you end up filling at \$100.25.
\end{keyterm}

\begin{keyterm}{Market Order}
An order that executes immediately at the best available prices. You prioritize speed over
price certainty.
\\\textbf{Example:} In a thin order book, a market buy may walk up the book.
\end{keyterm}

\begin{keyterm}{Limit Order}
An order that executes only at your specified price or better. You prioritize price certainty over
guaranteed fill.
\\\textbf{Example:} A buy limit at \$99.50 won't buy if the ask never reaches \$99.50.
\end{keyterm}

\begin{keyterm}{Tick}
The smallest allowed change in price (minimum price increment).
\\\textbf{Example:} If tick size is 0.01 USDT on a venue, prices move in discrete steps of 0.01.
\end{keyterm}

\subsection{Market Makers, Inventory, and Crypto Basis}

\begin{keyterm}{Inventory}
The net positions market makers accumulate while providing two-sided liquidity.
Carrying inventory creates risk if the market moves against the position.
\\\textbf{Example:} If a market maker has sold more than bought recently, they may be net short and
can lose if price rises.
\end{keyterm}

\begin{keyterm}{Skew}
An asymmetry in pricing/quoting (often referring to options implied-volatility skew, or more generally
that market makers price different strikes differently).
\\\textbf{Example:} Downside puts can have higher implied volatility, creating a skew across strikes.
\end{keyterm}

\begin{keyterm}{Spread Capture}
How market makers try to profit: repeatedly buying at the bid and selling at the ask.
If execution is mostly from passive flow, this can be a steady return stream.
\\\textbf{Example:} Earn small profits on many trades while keeping your quotes tight.
\end{keyterm}

\begin{keyterm}{Adverse Selection}
The risk that trades hitting your quotes are informed (you get ``picked off'') rather than random.
It can turn spread capture into losses.
\\\textbf{Example:} You quote tightly, but sophisticated traders know price is about to move and
your filled trades become systematically loss-making.
\end{keyterm}

\begin{keyterm}{Basis}
The price gap between two related instruments (commonly: futures/perps vs spot).
In crypto, basis is tightly connected to funding.
\\\textbf{Example:} If BTC perp trades \$200 above BTC spot, positive basis creates incentive for funding
to move the perp back toward spot.
\end{keyterm}

\begin{keyterm}{Open Interest}
The total number of outstanding derivative contracts (not yet closed). It reflects how much positioning
is currently ``hanging out'' in the system.
\\\textbf{Example:} Open interest rising alongside price can indicate growing leverage; falling open interest
can indicate deleveraging.
\end{keyterm}

% ────────────────────────────────────────────────────────────
\section{Quantitative and Algorithmic Trading}

\subsection{Systematic vs. Discretionary}
\begin{itemize}
  \item \textbf{Discretionary} — human makes all trading decisions. Can incorporate
        qualitative/contextual reasoning.
  \item \textbf{Systematic} — rules-based; computer executes all trades. Removes emotion.
        Requires robust backtesting and forward testing.
\end{itemize}

\subsection{Strategy Development Process}
\begin{enumerate}[label=\textbf{\arabic*.}]
  \item \textbf{Hypothesis} — define the edge you believe exists and why it should persist.
  \item \textbf{Data collection} — quality historical data is non-negotiable.
  \item \textbf{Backtest} — simulate strategy on historical data. Watch for overfitting.
  \item \textbf{Walk-forward test} — test on data not used in development (out-of-sample).
  \item \textbf{Paper trading} — simulate live with real-time data, no real money.
  \item \textbf{Live trading (small size)} — deploy with minimal capital first.
  \item \textbf{Scale} — increase size only after demonstrable live performance.
\end{enumerate}

\begin{warningbox}
\textbf{Overfitting} is the cardinal sin of backtesting: a strategy that perfectly fits
historical data but fails on new data. The more parameters you optimise, the more you
risk fitting noise. Always hold out out-of-sample data for validation.
\end{warningbox}

\subsection{Common Quant Strategies}

\subsubsection{Market Making (Spread Capture)}

A market maker acts as a \textbf{liquidity provider} by continuously quoting both a bid
and an ask. The goal is not to predict direction, but to capture the \textbf{spread}
between the two sides as other participants trade against those quotes.

\begin{keyterm}{Fair Value Estimation}
You must have a strong estimate of the asset's ``true'' price. All quotes are placed
symmetrically (or asymmetrically) around this fair value.
\\\textbf{Example:} If fair value is 10{,}000, you might bid 9{,}998 and ask 10{,}002,
capturing the 4-point spread on each round-trip.
\end{keyterm}

\begin{keyterm}{Inventory Risk \& Skewing}
The danger of accumulating too much of an asset right as its fair value crashes.
To mitigate, market makers \textbf{skew} their quotes based on current position:
\begin{itemize}[noitemsep]
  \item \textbf{Long inventory} (holding too much): lower the ask to sell faster;
        lower the bid to stop buying more.
  \item \textbf{Short inventory} (negative position): raise the bid to buy faster;
        raise the ask to stop selling more.
\end{itemize}
This dynamically rebalances the book without taking directional risk.
\end{keyterm}

\begin{warningbox}
\textbf{Market making requires a mean-reverting asset.}
If the underlying price trends strongly, the market maker gets run over:
buy at 9{,}999, price drops to 9{,}900, and you are holding a depreciating
inventory with no exit at fair value. Pure MM should only be applied to assets
that oscillate around a stable fair value. For noisier or trending assets,
momentum gating (see below) is required to shut off quoting when price moves
persistently in one direction.
\end{warningbox}

\subsubsection{Arbitrage}

Exploiting a definitive mispricing for a risk-free or highly probable profit, when the
market price momentarily violates a mathematical or structural boundary.

\begin{itemize}
  \item \textbf{Absolute / Fair Value Arb:} If an asset has a locked fair value
        (e.g., a pegged stablecoin), crossing the spread to buy at 9{,}999 or sell at
        10{,}001 guarantees a profit. Executed aggressively via market orders.
  \item \textbf{Aggressive Arb (Quote Snipe):} A specific execution mode where the
        algorithm scans the live order book and \emph{hits} any ask priced below fair
        value or any bid priced above fair value. Rather than passively waiting for
        others to trade against your quotes, the algorithm aggressively takes
        mispriced orders the moment they appear. No inventory is accumulated; each
        fill is immediately profitable relative to fair value.
  \item \textbf{Statistical Arbitrage (Pairs Trading):} Two highly correlated assets
        (e.g., two competing stocks in the same sector). If their price spread widens
        beyond historical norms, short the overperformer and long the underperformer,
        betting the spread will converge. The core assumption is mean reversion of the
        \emph{pair}, not of either asset individually.
\end{itemize}

\begin{warningbox}
Arb opportunities are extremely short-lived in liquid markets — competing algorithms
close them within milliseconds. Execution speed and transaction costs determine
whether a theoretical arb is actually profitable.
\end{warningbox}

\subsubsection{Mean Reversion}

The statistical assumption that an asset's price will eventually return to its
long-term average. Prices oscillate due to temporary imbalances in supply and demand;
the strategy profits by fading those temporary extremes.

\begin{center}
\begin{tabular}{>{\bfseries}p{3.5cm} p{9cm}}
\toprule
Indicator & How It Is Used \\
\midrule
SMA / EMA & Rolling average acts as the dynamic ``fair value''. EMA reacts faster
            to recent changes than SMA. Buy when price falls well below; sell when
            it rises well above. The EMA update rule is:
            $\text{EMA}_t = \alpha \cdot P_t + (1 - \alpha) \cdot \text{EMA}_{t-1}$,
            where $\alpha \in (0,1)$ is the decay factor. A common choice is
            $\alpha = 0.15$, giving the most recent price roughly 13\% weight and
            decaying older prices exponentially. Smaller $\alpha$ = smoother,
            slower-reacting average. \\
Bollinger Bands & Upper and lower bands set at $\pm N$ standard deviations from the
                  moving average. Buy at the lower band; sell at the upper band.
                  Automatically adjusts thresholds for current volatility. \\
VWAP & Volume Weighted Average Price: average price weighted by volume throughout
       the session. Institutional benchmark for fair intraday execution; acts as
       a mean-reversion anchor intraday. \\
\bottomrule
\end{tabular}
\end{center}

\begin{notebox}
\textbf{Aggressive Take:} If price spikes dramatically away from the EMA, an algorithm
will aggressively cross the spread (market order) to ``fade'' the spike, betting on a
fast snapback. This is higher risk but captures more of the reversion premium.
\end{notebox}

\subsubsection{Momentum and Trend Following}

The opposite of mean reversion. Momentum assumes that assets in motion tend to stay
in motion: a breakout upward is bet to continue before any reversal occurs.
CTA-style trend following applies this across all asset classes.

\begin{keyterm}{RSI (Relative Strength Index)}
An oscillator on a 0--100 scale measuring the speed and magnitude of recent price
moves. $>70$ is conventionally ``overbought'' (strong upward momentum); $<30$ is
``oversold'' (strong downward momentum). Used both as a signal and as a gate.
\end{keyterm}

\begin{keyterm}{Momentum Gating (Hybrid Use)}
Momentum indicators are frequently used \emph{defensively} inside mean-reversion or
market-making strategies as a gate to avoid catching a falling knife:
\begin{itemize}[noitemsep]
  \item \textbf{Without a gate:} Price drops sharply; mean-reversion algorithm
        buys the ``dip''; price keeps falling; algorithm averages down; permanent
        trend destroys the position.
  \item \textbf{With a momentum gate:} Price drops sharply. Algorithm wants to buy,
        but checks RSI. If RSI is at 20 (extreme downward momentum), the algorithm
        \emph{refuses} to enter until momentum slows, avoiding the trend entirely.
\end{itemize}
\end{keyterm}

\begin{itemize}
  \item \textbf{Factor investing} — systematic exposure to well-documented return
        premia: value, quality, size, low-volatility, and momentum factors.
\end{itemize}

\begin{notebox}
\textbf{Asset Behaviour $\to$ Strategy Mapping.}
The right strategy depends entirely on how the asset's price \emph{behaves},
not on preference.

\begin{center}
\begin{tabular}{>{\bfseries}p{4cm} p{3.5cm} p{5.5cm}}
\toprule
Price Behaviour & Right Strategy & Why \\
\midrule
Stable, mean-reverts to known fair value &
  Pure Market Making &
  Earn the spread repeatedly; inventory risk is low because price returns to fair value \\
Noisy, can trend temporarily &
  Mean Reversion + Momentum Gate &
  Fade extremes, but halt if RSI signals a sustained trend to avoid catching a falling knife \\
Persistent trend &
  Momentum / Trend Following &
  MM and mean-reversion both lose against a trend; ride the direction instead \\
Two correlated assets diverge &
  Stat Arb / Pairs Trading &
  Short the outperformer, long the underperformer; profit from convergence \\
Definitive mispricing vs.\ fair value &
  Aggressive Arb &
  Hit the mispriced order immediately; no passive quoting required \\
\bottomrule
\end{tabular}
\end{center}
\end{notebox}

% ────────────────────────────────────────────────────────────
\subsection{Algorithmic Risk Management}

Beyond placing trades, algorithms must protect capital \emph{systematically}. The
following are standard safeguards implemented at the strategy level.

\begin{keyterm}{Adverse Selection Protection}
The risk that the counterparty taking your market-making quote is an informed trader
who knows something you don't. If your bids are filled rapidly across a very short
time window, the market is likely crashing.
Algorithms detect this pattern and \textbf{pull remaining quotes immediately} rather
than continuing to provide liquidity into a fast-moving adverse move.
\end{keyterm}

\begin{keyterm}{Time-of-Day Adaptation}
Volatility and volume are not uniform across the trading session. Markets typically
exhibit high volatility at the open and close, and lower volatility mid-session.
Algorithms adjust spread thickness, position sizing, and participation rate based
on the time of day to avoid being caught in abnormally wide moves.
\end{keyterm}

\begin{keyterm}{Drawdown Guards (Kill Switches)}
Hardcoded circuit breakers that halt the algorithm entirely if realised or unrealised
P\&L drops below a threshold. Critical for preventing runaway losses from logic bugs,
feed issues, or extreme market events — the algorithm simply stops trading until a
human reviews.
\end{keyterm}

\begin{keyterm}{Position Unwinding (Flattening)}
The deliberate closing of all open positions before a predetermined time (e.g., market
close), to avoid holding inventory through periods when the algorithm cannot react —
overnight news gaps, weekend risk, or exchange maintenance windows.
``Going flat'' at end-of-day is standard practice for intraday systematic strategies.
\end{keyterm}

\begin{keyterm}{State Persistence Between Ticks}
In tick-driven or event-driven algorithms, each execution cycle may be stateless by
default — the algorithm is called, runs, and exits, with no memory of previous ticks.
To maintain rolling indicators (EMA, RSI, VWAP accumulators, inventory counts),
the algorithm must \textbf{serialise its state} at the end of each cycle and
\textbf{deserialise} it at the start of the next.
\\\textbf{Common pattern:} Encode the state dictionary as a JSON string and write it
to a persistent field (e.g., \texttt{traderData} in competition platforms). On the
next tick, parse the JSON string to restore the previous EMA value, position size,
and any other running variables before computing new signals.
\end{keyterm}

% ────────────────────────────────────────────────────────────
\section{Behavioural Finance}

\begin{keyterm}{Behavioural Finance}
The study of how cognitive biases and emotional responses cause investors to make
irrational decisions, leading to systematic market mispricings.
\end{keyterm}

\begin{center}
\begin{tabular}{>{\bfseries}p{4cm} p{9cm}}
\toprule
Bias & Description \\
\midrule
Loss Aversion & Losses feel roughly $2\times$ worse than equivalent gains feel good \\
Overconfidence & Overestimating one's predictive accuracy and skill \\
Anchoring & Over-relying on the first piece of information seen (e.g., purchase price) \\
Recency Bias & Overweighting recent events; extrapolating the current trend \\
Herd Behaviour & Following the crowd; contributing to bubbles and crashes \\
Disposition Effect & Holding losers too long, selling winners too early \\
Sunk Cost Fallacy & Holding a bad investment to ``get back to even'' \\
Narrative Fallacy & Constructing plausible stories to explain random price moves \\
\bottomrule
\end{tabular}
\end{center}

% ────────────────────────────────────────────────────────────
\section{Recommended Learning Roadmap}

\subsection{Beginner (Months 1--3)}
\begin{itemize}[noitemsep]
  \item Complete this entire notes document
  \item Read: \textit{The Intelligent Investor} — Benjamin Graham
  \item Read: \textit{A Random Walk Down Wall Street} — Burton Malkiel
  \item Open a \textbf{paper trading account} (Interactive Brokers, Thinkorswim)
  \item Study and annotate 100 historical charts
\end{itemize}

\subsection{Intermediate (Months 4--9)}
\begin{itemize}[noitemsep]
  \item Read: \textit{Market Wizards} — Jack Schwager
  \item Read: \textit{Trading in the Zone} — Mark Douglas (psychology)
  \item Read: \textit{Technical Analysis of the Financial Markets} — John Murphy
  \item Develop and paper trade a \emph{single, well-defined} strategy for 3 months minimum
  \item Begin tracking trades in a \textbf{trading journal}
\end{itemize}

\subsection{Advanced (Month 10+)}
\begin{itemize}[noitemsep]
  \item Trade live, minimum size; focus on process, not P\&L
  \item Read: \textit{Reminiscences of a Stock Operator} — Edwin Lefèvre
  \item Study options via tastytrade or Options Education (CBOE)
  \item Explore macro with: \textit{The Alchemy of Finance} — George Soros
  \item Consider Python for backtesting: \texttt{pandas}, \texttt{vectorbt}, \texttt{backtrader}
\end{itemize}

% ────────────────────────────────────────────────────────────
\section{Essential Tools and Resources}

\begin{center}
\begin{tabular}{>{\bfseries}p{3.5cm} p{3cm} p{6.5cm}}
\toprule
Tool & Category & Use Case \\
\midrule
TradingView & Charting & Best charting platform; scripting via Pine Script \\
Interactive Brokers & Broker & Best for serious traders; global access, low fees \\
Bloomberg Terminal & Data/News & Institutional standard (expensive); use free alternatives \\
Yahoo Finance & Data & Free fundamental data and price history \\
EDGAR (SEC) & Filings & Official source for US company filings \\
Finviz & Screener & Stock screening by technical and fundamental filters \\
Quandl / FRED & Economic Data & Macro and economic time series data \\
Python / R & Quant & Backtesting, data analysis, strategy automation \\
\bottomrule
\end{tabular}
\end{center}

% ────────────────────────────────────────────────────────────
\section{Final Principles to Internalise}

\begin{enumerate}[label=\textbf{\Roman*.}]
  \item \textbf{Capital preservation first}. You cannot profit if you blow up.
  \item \textbf{Process over outcome}. A losing trade within your rules is fine.
        A winning trade outside your rules is dangerous.
  \item \textbf{The market is always right}. Your opinion is irrelevant; price is fact.
  \item \textbf{Cut losers, let winners run}. This is the opposite of what emotions dictate.
  \item \textbf{There is no holy grail}. Any edge erodes over time; keep learning.
  \item \textbf{Consistency beats brilliance}. A 60\% win rate with 2:1 RRR,
        repeated 1{,}000 times, compounds to large wealth.
  \item \textbf{Know your edge}. If you cannot articulate why your strategy works,
        you do not have one.
  \item \textbf{Keep a trading journal}. The data in your own trade history is the
        most valuable dataset you will ever have.
\end{enumerate}

\vspace{2em}
\begin{tcolorbox}[colback=lightblue, colframe=darkblue]
\centering
\textbf{``The goal of a successful trader is to make the best trades.\\
Money is secondary.''} — Alexander Elder
\end{tcolorbox}
